\pkg{cpop}: Detecting changes in piecewise-linear signals

### 2022-11-07 latex/article.tex
Add files via upload

\subsection{R Packages for Changepoint Detection}

There are both many different types of change that one may wish to detect, and many different approaches to detecting multiple changes. Consequently there are a wide range of change algorithms with associated packages in \proglang{R}. For example the \pkg{changepoint} package \citep{killick2014changepoint} implements dynamic programming algorithms, such as PELT, for detecting changes in mean, variance or mean and variance. Other dynamic programming algorithms include \pkg{fpop} and \pkg{gfpop} \citep{runge2020gfpop} that implements the functional optimal partitioning algorithm \citep{maidstone2017optimal} for detecting changes in mean, with the latter package allowing for flexibility as to how the mean changes (such as monotonically increasing) and for different loss functions for measuring fit to data. The \pkg{breakfast} package implements a range of methods based on recursively applying a test to detect a single change, for example wild binary segmentation \citep{fryzlewicz2014wild} and IsolateDetect \citep{anastasiou2022detecting}; whilst \pkg{mosum} \citep{meier2021mosum} implements the MOSUM procedure. Packages \pkg{stepR} and \pkg{FDRseg} implement the multiscale approaches of \cite{frick2014multiscale}, \cite{pein2017heterogeneous} and \cite{li2016fdr}.

@@

However, as mentioned above, there are more limited methods for specifically detecting changes-in-slope. The trend filtering algorithm can be implemented using the \code{trendfilter} function from the \pkg{genlasso} \citep{genlasso} package, and the NOT algorithm can be implemented using the \pkg{not} package or is available within \pkg{breakfast}. However current implementations of these do not allow for unevenly spaced, heterogeneous observations or minimum segment lengths. (Though there is flexibility within the \pkg{genlasso} package for implementing general lasso algorithms, and these can be constructed to fit a trend-filtering model to unevenly spaced data).

@@

The following code demonstrates the \code{simchangeslope} function and displays the data along with the (true) line segments and the locations of the changes in slope (see Figure \ref{fig:simulate-example}).
\begin{CodeChunk}
\begin{CodeInput}
R> library("cpop")
R> library("ggplot2")
R> changepoints <- c(0, 25, 50, 100)
R> change.slope <- c(0.2, -0.3, 0.2, -0.1)
R> x <- 1:200
R> sd <- 0.8
R> y <- simchangeslope(x, changepoints, change.slope, sd)
R> df <- data.frame("x" = x, "y" = y)
R> p <- ggplot(data = df, aes(x = x, y = y))
R> p <- p + geom_point(alpha = 0.4)
R> p <- p + geom_vline(xintercept = changepoints,
+                      color = "red",
+                      linetype = "dashed")
R> mu <- simchangeslope(x, changepoints, change.slope, sd = 0)
R> p <- p + geom_line(aes(y = mu), color = "blue")
R> p <- p + theme_bw()
R> print(p)
\end{CodeInput}
\end{CodeChunk}
%
%
\begin{figure}
\centering
%\includegraphics[scale=0.5]{figures/simulate-example.pdf}
\includegraphics[width=0.6\linewidth]{figures/simulate_example_ggplot.pdf}
\caption{Simulated data (black dots) with true mean (blue dashed line) and changepoints (vertical red dashed lines).}
\label{fig:simulate-example}
\end{figure}
%
%
\subsection{Determining Changes in Slope}
%
%
The function \code{cpop} is used to determine the locations of changes in slope.
%
%
\begin{CodeInput}
cpop(y, x = 1:length(y) - 1, grid = x, beta = 2 * log(length(y)), 
    sd = 1, minseglen = 0, prune.approx = FALSE)
\end{CodeInput}
It takes the following arguments :
%
%
\begin{itemize}
 \item  \code{y} - A vector of length $n$ containing the data.
\item \code{x} -  A vector of length $n$ containing the times/locations of data points. Default value is NULL, in which case the locations are set to be $0,1,\ldots,n-1$, corresponding to evenly spaced data.
\item \code{grid} -  An ordered vector of possible locations for the change points. If this is NULL, then this is set to $x$, the vector of times/locations of the data points.
\item \code{beta} -  A positive real value for the penalty, $\beta$ in (\ref{eq:penalisedcost2}), incurred for adding a changepoint. The larger the penalty, the fewer changepoints will be detected.
\item \code{sd} - Estimate of residual standard deviation. Can be a single numerical value if it is the same for all data points, or a vector of $n$ values for the case of varying standard deviation. 
\item \code{minseglen} -  The minimum allowable segment length, i.e. distance between successive changepoints. The default is that no minimum segment length is imposed.
\item \code{prune.approx} -  Only relevant if a minimum segment length is set. If True, \code{cpop} will use an approximate pruning algorithm that will speed up computation but may occasionally lead to a sub-optimal solution in terms of the estimated changepoint locations. If the minimum segment length is 0, then an exact pruning algorithm is possible and is used.
\end{itemize}
The \code{cpop} function returns an S4 object for which a number of generic methods, including \code{plot} and \code{summary}, are provided. 
%
%
%\subsubsection{simulated data}
The following demonstrates how the \code{cpop} function can be used to determine changes in slope, by analysing the data we simulated and plotted in Figure \ref{fig:simulate-example}. It uses the default penalty, $\beta=2\log n$, and we assume that the true noise standard deviation, 0.8, is known. The \code{summary} function is used to provide an overview of the analysis parameters along with estimated changepoint locations, corresponding fitted line segments, and the (weighted) residual sum of squares (RSS) for each segment. 
%
%
\begin{CodeChunk}
\begin{CodeInput}
R> res <- cpop(y, x, sd = 0.8)
R> summary(res)
\end{CodeInput}
\begin{CodeOutput}
cpop analysis with n = 200 and penalty (beta)  = 10.59663

@@

To analyse the data we use \code{cpop} as before. (The only difference is that for evenly spaced data one can omit the \code{x} argument -- but it must be included for unevenly spaced data.)
\begin{CodeChunk}
\begin{CodeInput}
R> res <- cpop(y, x, sd = 0.8)
\end{CodeInput}
\end{CodeChunk}
%
%
Figure \ref{fig:cpop-example-uneven} shows a plot of the simulated data and estimated changepoints and mean function.

@@

It is interesting to compare two estimates of the changepoints, one where we assume a fixed noise standard deviation, and one where we assume the true noise standard deviation. For the former it is natural to set this value so the the average variance of the noise is correct. 
\begin{CodeChunk}
\begin{CodeInput}
R> res <- cpop(y, x, sd = sqrt(mean(sd^2)))
R> res.true <- cpop(y, x, sd = sd)
\end{CodeInput}
\end{CodeChunk}
Here \code{res} contains the results where we assume a fixed noise standard deviation, and \code{res.true} where we use the true values. Figure \ref{fig:cpop-example-heterogeneous} shows the results -- and we can see that wrongly assuming homogeneous noise leads to detecting two false positive changepoints in regions where the noise variance is above what was assumed.

@@

We use a smaller value for the penalty due to the smaller grid size, and the fact that this is a preliminary step to find roughly where the changes are: so the key is to avoid missing changes. Spurious changes can still be removed when we perform our final run of \code{cpop}. 
\begin{CodeChunk}
\begin{CodeInput}
R> res.coarse <- cpop(y, x, grid=1:399*16, beta = 2 * log(400))
\end{CodeInput}
\end{CodeChunk}

@@

R> res.fine <- cpop(y, x, grid, beta = 2 * log(length(x)))
\end{CodeInput}
\end{CodeChunk}

@@

We then estimated the changepoint locations both without a minimum segment length, and with minimum segment lengths of 10, 30 and 40. To run \code{cpop} with a minimum segment length of 10:
\begin{CodeChunk}
\begin{CodeInput}
R> res.min <- cpop( y, x, beta = 2*( log( length(y) ) ),
  +. minseglen = 10, sd = sqrt(2) )
\end{CodeInput}
\end{CodeChunk}
The argument \code{sd = sqrt(2)} is because $t_4$ distributed noise has a variance of 2.
%
%\begin{figure}
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg1.pdf}}
%\end{minipage}%
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg2.pdf}}
%\end{minipage}%
%\newline
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg3.pdf}}
%\end{minipage}%
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg4.pdf}}
%\end{minipage}%
%\centering
%\includegraphics{figures/cpop-example-uneven.pdf}
%\caption{Results of analysing data with $t_4$ noise with no minimum segment length (top left) and minimum segment lengths of 10 (top right), 30 (bottom left) and 40 (bottom right). In each plot we show data (grey dots), true mean (blue dashed line), true changepoints (blue vertical dashed lines), estimated mean (black line) and estimated changepoints (red vertical liens)}
%\label{fig:cpop-minseg}
%\end{figure}
%
%
\begin{figure}
\centering
{\includegraphics[scale=.5]{figures/cpop_minseg_ggplot.pdf}}
\caption{Results of analysing data with $t_4$ noise with no minimum segment length (top left) and minimum segment lengths of 10 (top right), 30 (bottom left) and 40 (bottom right). In each plot we show data (grey dots), true mean (blue dashed line), true changepoints (blue vertical dashed lines), estimated mean (black line) and estimated changepoints (red vertical liens)}
\label{fig:cpop-minseg}
\end{figure}
%
%
Results from CPOP with different minimum segment lengths are shown in Figure \ref{fig:cpop-minseg}. If we do not impose a minimum segment length, then we estimate 11 changepoints, including three cluster of changes that overfit to the noise. By imposing a minimum segment length of 10 we avoid the over-fitting. For this example, the computational cost of running \code{cpop} with the minimum segment length is about 15 times larger than when we do not assume a minimum segment length.

@@

The idea of CROPS is that it allows a penalised cost method to be implemented for all penalty values in an interval. This is implemented within the \pkg{cpop} package by the function:
\begin{CodeInput}
cpop.crops(y, x = 1:length(y), grid = x, beta_min = 1.5 * log(length(y)),
  beta_max = 2.5 * log(length(y)), sd = 1, minseglen = 0, prune.approx = FALSE)
\end{CodeInput}
The arguments of \code{cpop.crops} are identical to those of \code{cpop} except that, rather than specifying a single penalty value (\code{beta}), the range of penalty values to be used is specified by \code{beta_min} and \code{beta_max}, which fix the smallest and largest penalty value to be used. The output is an instance of an S4 class that contains details of all segmentations found by minimising the penalised cost for some penalty value in the interval between \code{beta_min} and \code{beta_max}.

@@

We simulated data as per Section \ref{sec:cpop} but with \code{sd = 1.5}. To run CPOP for a range of $\beta\in[7.5, 50]$ we use
\begin{CodeInput}
R> res.crops <- cpop.crops(y, x, beta_min = 5, beta_max = 50)
\end{CodeInput}
For our example $2 \log n = 10.6$, so this is equivalent to trying noise standard deviation in the range $[0.8,2.2]$.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> res.crops <- cpop.crops(y, x, beta_min = 8, beta_max = 200)
R> segs <- segmentations(res.crops)
R> p <- ggplot(data = segs, aes(x = m))
R> p <- p + geom_line(aes(y = Qm))
R> p <- p + geom_vline(xintercept = 10, color = "red")
R> p <- p + xlab("No. of changepoints") + ylab("unpenalised cost")
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
This runs \code{cpop.crops} and then uses the fact that the output of \code{segmentations} includes columns that give the number of changepoints and the unpenalised cost of each segmentation. These columns are labelled \code{"m"} and \code{"Qm"} respectively.
The plot gives a clear elbow, see Figure \ref{fig:crops.cpopB}, and this corresponds to a correct estimate of the number of changes. %The fitted changes and estimated mean function are also shown in Figure \ref{fig:crops.cpopB}.
%
%\begin{figure}
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-cropsB1.pdf}}
%\end{minipage}%
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-cropsB2.pdf}}
%\end{minipage}%
%\newline
%\begin{minipage}{1.0\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-cropsB3.pdf}}
%\end{minipage}%
%\caption{Correlated noise example. Output from \code{cpop.crops} (top left), unpenalised cost against number of changepoint (rop right) and estimate from segmentation corresponding to "elbow" (bottom).}
%\label{fig:crops.cpopB}
%\end{figure}
%
\begin{figure}
\centering
{\includegraphics{figures/cpop_crops_ggplot.pdf}}
\caption{Correlated noise example. Output from \code{cpop.crops} (top left), unpenalised cost against number of changepoint (top right) and estimate from segmentation corresponding to "elbow" (bottom).}
\label{fig:crops.cpopB}
\end{figure}

@@

R> res <- cpop(y, x, grid, sd = sqrt(sig2), minseg = 0.2, beta = 2*log(200))
\end{CodeInput}
\end{CodeChunk}
Here we estimate the noise variance, \code{sig2}, based on the variance of the double difference of the data. For regions where the mean is linear and the data is evenly spaced, taking the double difference will lead to a mean zero process. If the noise is IID with variance $\sigma^2$ then the double-differenced process will have a marginal variance that is $6\sigma^2$. Thus our estimate is the empirical mean of the square of the double-difference data divided by 6. The original data is evenly spaced in in terms of wavenumber, but as we take logs, \code{x} is unevenly spaced: so this estimator will be biased in our setting. However it will give a reasonable ball-park figure for an initial run of \code{cpop} -- the residuals from which can then be used to get a better estimate of the noise variance. We use a evenly spaced grid for possible change-point locations. To avoid the potential for adding multiple changepoints between two observations, we set a minimum segment length of 0.09 (as the largest distance between consecutive \code{x} values is 0.08).

@@

The output is shown in the top-right plot of Figure \ref{fig:crops.data}, and appears to be over-fitting to the early part of the series. This is because the noise variance is heterogeneous, and decreasing with \code{x}. However, given our initial fit we can use the residuals to estimate the noise variance. The noise for the spectra is expected to be approximately inversely proportional to the wavenumber. By using a Taylor-expansion, we have the variance of the noise for the log-spectra should be approximately the variance of the noise of the spectra divided by the square of the mean of the spectra. As the mean of the spectra is roughly a power of the wavenumber, this suggests using a model for the variance, $\sigma^2_x$ say, depending on $x$ as $\log \sigma^2_x = a + bx$: so that the variance is proportional to some power of the wavenumber. We can estimate the parameters of this model by maximising the log-likelihood of Gaussian model for the residuals with this form for the variance.
\begin{CodeChunk}
\begin{CodeInput}
R> r2 <- residuals(res)^2
R> loglik <- function(par)
R> { 
R>  return(length(r2) * par[1] + par[2] * sum(x) 
  + sum(r2 / (exp(par[1]+par[2]*x))))
R> }
R> est.hat <- optim(c(0,0), loglik)
R> sig2 <- exp(est.hat$par[1] + est.hat$par[2]*x)
R> res2 <- cpop(y, x, grid, sd = sqrt(sig2), 
+               minseg = 0.2, beta =2 * log(200))
\end{CodeInput}
\end{CodeChunk}
Here we have calculated the maximum likelihood estimates by using \code{optim} to minimise minus the log-likelihood. The resulting output from \code{cpop} is shown in the bottom left plot of Figure \ref{fig:crops.data}.  The first two changes could represent real regime transitions relating to the inviscid fluid physics that one would see in the real ocean \cite[see Figure 6a of][]{callies2013interpreting}, while the three changes for the largest values of \code{x} may relate to a breakdown in the numerical ocean model near the highest wavenumber of the ocean model grid \cite[]{soufflet2016effective}. The estimates of the spectra for all four series, obtained by repeating this approach, is also shown in Figure \ref{fig:crops.data}. For this application, the residuals from the fitted model appear to be uncorrelated and sub-Gaussian, so using the fit based on the default penalty choice is reasonable. Though one could also explore segmentations for other penalty choices using \code{cpop.crops} as in the previous section.

>>>>

\subsection{Software for Changepoint Detection}

Currently most software for changepoint detection is available in \proglang{R} \cite[the main exception being the \pkg{ruptures} \proglang{Python} package of][which has similar functionality to the \pkg{changepoint } described below]{truong2018ruptures}.
There are both many different types of change that one may wish to detect, and many different approaches to detecting multiple changes. Consequently there are a wide range of change algorithms with associated packages in \proglang{R}. For example the \pkg{changepoint} package \citep{killick2014changepoint} implements dynamic programming algorithms, such as PELT, for detecting changes in mean, variance or mean and variance. Other dynamic programming algorithms include \pkg{fpop} and \pkg{gfpop} \citep{runge2020gfpop} that implements the functional optimal partitioning algorithm \citep{maidstone2017optimal} for detecting changes in mean, with the latter package allowing for flexibility as to how the mean changes (such as monotonically increasing) and for different loss functions for measuring fit to data. The \pkg{breakfast} package implements a range of methods based on recursively applying a test to detect a single change, for example wild binary segmentation \citep{fryzlewicz2014wild} and IsolateDetect \citep{anastasiou2022detecting}; whilst \pkg{mosum} \citep{meier2021mosum} implements the MOSUM procedure. Packages \pkg{stepR} and \pkg{FDRseg} implement the multiscale approaches of \cite{frick2014multiscale}, \cite{pein2017heterogeneous} and \cite{li2016fdr}.

@@

However, as mentioned above, there are more limited methods for specifically detecting changes-in-slope.  The trend filtering algorithm can be implemented using the \code{trendfilter} function from the \pkg{genlasso} \citep{genlasso} package, and the NOT algorithm can be implemented using the \pkg{not} package or is available within \pkg{breakfast}. However current implementations of these do not allow for unevenly spaced, heterogeneous observations or minimum segment lengths, which are all features that can be included within the latest release of the \pkg{cpop} package that is described in this article. (Though there is flexibility within the \pkg{genlasso} package for implementing general lasso algorithms, and these can be constructed to fit a trend-filtering model to unevenly spaced data).

@@

The following code demonstrates the \code{simchangeslope} function and displays the data along with the (true) line segments and the locations of the changes in slope (see Figure \ref{fig:simulate-example}).
\begin{CodeChunk}
\begin{CodeInput}
R> library("cpop")
R> library("ggplot2")
R> changepoints <- c(0, 25, 50, 100)
R> change.slope <- c(0.2, -0.3, 0.2, -0.1)
R> x <- 1:200
R> sd <- 0.8
R> y <- simchangeslope(x, changepoints, change.slope, sd)
R> df <- data.frame("x" = x, "y" = y)
R> p <- ggplot(data = df, aes(x = x, y = y))
R> p <- p + geom_point(alpha = 0.4)
R> p <- p + geom_vline(xintercept = changepoints,
+                      color = "red",
+                      linetype = "dashed")
R> mu <- simchangeslope(x, changepoints, change.slope, sd = 0)
R> p <- p + geom_line(aes(y = mu), color = "blue")
R> p <- p + theme_bw()
R> print(p)
\end{CodeInput}
\end{CodeChunk}
%
%
\begin{figure}
\centering
%\includegraphics[scale=0.5]{figures/simulate-example.pdf}
\includegraphics[width=0.6\linewidth]{figures/simulate_example_ggplot.pdf}
\caption{Simulated data (black dots) with true mean (blue dashed line) and changepoints (vertical red dashed lines).}
\label{fig:simulate-example}
\end{figure}
%
%
\subsection{Determining Changes in Slope}
%
%
The function \code{cpop} is used to determine the locations of changes in slope.
%
%
\begin{CodeInput}
cpop(y, x = 1:length(y) - 1, grid = x, beta = 2*log(length(y)), 
    sd =  sqrt(mean(diff(diff(y))^2)/6), minseglen = 0, 
    prune.approx = FALSE)
\end{CodeInput}
It takes the following arguments :
%
%
\begin{itemize}
 \item  \code{y} - A vector of length $n$ containing the data.
\item \code{x} -  A vector of length $n$ containing the times/locations of data points. Default value is NULL, in which case the locations are set to be $0,1,\ldots,n-1$, corresponding to evenly spaced data.
\item \code{grid} -  An ordered vector of possible locations for the change points. If this is NULL, then this is set to $x$, the vector of times/locations of the data points.
\item \code{beta} -  A positive real value for the penalty, $\beta$ in (\ref{eq:penalisedcost2}), incurred for adding a changepoint. The larger the penalty, the fewer changepoints will be detected. The default value is \code{beta = 2*log(length(y))}.
\item \code{sd} - Estimate of residual standard deviation. Can be a single numerical value if it is the same for all data points, or a vector of $n$ values for the case of varying standard deviation. The default value is \code{sd = sqrt(mean(diff(diff(y))**2)/6)}.
\item \code{minseglen} -  The minimum allowable segment length, i.e. distance between successive changepoints. The default is that no minimum segment length is imposed.
\item \code{prune.approx} -  Only relevant if a minimum segment length is set. If True, \code{cpop} will use an approximate pruning algorithm that will speed up computation but may occasionally lead to a sub-optimal solution in terms of the estimated changepoint locations. If the minimum segment length is 0, then an exact pruning algorithm is possible and is used.
\end{itemize}
The \code{cpop} function returns an S4 object for which a number of generic methods, including \code{plot} and \code{summary}, are provided. 
%
%
%\subsubsection{simulated data}
The following demonstrates how the \code{cpop} function can be used to determine changes in slope, by analysing the data we simulated and plotted in Figure \ref{fig:simulate-example}. It uses the default penalty, $\beta=2\log n$, and we assume that the true noise standard deviation, 0.8, is known. The \code{summary} function is used to provide an overview of the analysis parameters along with estimated changepoint locations, corresponding fitted line segments, and the (weighted) residual sum of squares (RSS) for each segment. 
%
%
\begin{CodeChunk}
\begin{CodeInput}
R> res <- cpop(y, x, beta = 2*log(length(y)), sd = 0.8)
R> summary(res)
\end{CodeInput}
\begin{CodeOutput}
cpop analysis with n = 200 and penalty (beta)  = 10.59663

@@

To analyse the data we use \code{cpop} as before. (The only difference is that for evenly spaced data one can omit the \code{x} argument -- but it must be included for unevenly spaced data.)
\begin{CodeChunk}
\begin{CodeInput}
R> res <- cpop(y, x, beta = 2*log(length(y)), sd = 0.8)
\end{CodeInput}
\end{CodeChunk}
%
%
Figure \ref{fig:cpop-example-uneven} shows a plot of the simulated data and estimated changepoints and mean function.

@@

It is interesting to compare two estimates of the changepoints, one where we assume a fixed noise standard deviation, and one where we assume the true noise standard deviation. For the former it is natural to set this value so the the average variance of the noise is correct. 
\begin{CodeChunk}
\begin{CodeInput}
R> res <- cpop(y, x, beta = 2*log(length(y)), sd = sqrt(mean(sd^2)))
R> res.true <- cpop(y, x, beta = 2*log(length(y)), sd = sd)
\end{CodeInput}
\end{CodeChunk}
Here \code{res} contains the results where we assume a fixed noise standard deviation, and \code{res.true} where we use the true values. Figure \ref{fig:cpop-example-heterogeneous} shows the results -- and we can see that wrongly assuming homogeneous noise leads to detecting two false positive changepoints in regions where the noise variance is above what was assumed.

@@

We use a smaller value for the penalty due to the smaller grid size, and the fact that this is a preliminary step to find roughly where the changes are: so the key is to avoid missing changes. Spurious changes can still be removed when we perform our final run of \code{cpop}. 
\begin{CodeChunk}
\begin{CodeInput}
R> res.coarse <- cpop(y, x, grid=1:399*16, beta = 2*log(400), sd = 1)
\end{CodeInput}
\end{CodeChunk}

@@

R> res.fine <- cpop(y, x, grid, beta = 2*log(length(x)), sd = 1)
\end{CodeInput}
\end{CodeChunk}

@@

We then estimated the changepoint locations both without a minimum segment length, and with minimum segment lengths of 10, 30 and 40. To run \code{cpop} with a minimum segment length of 10:
\begin{CodeChunk}
\begin{CodeInput}
R> res.min <- cpop(y, x, beta = 2*log(length(y)),
  +. minseglen = 10, sd = sqrt(2) )
\end{CodeInput}
\end{CodeChunk}
The argument \code{sd = sqrt(2)} is because $t_4$ distributed noise has a variance of 2.
%
%\begin{figure}
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg1.pdf}}
%\end{minipage}%
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg2.pdf}}
%\end{minipage}%
%\newline
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg3.pdf}}
%\end{minipage}%
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-minseg4.pdf}}
%\end{minipage}%
%\centering
%\includegraphics{figures/cpop-example-uneven.pdf}
%\caption{Results of analysing data with $t_4$ noise with no minimum segment length (top left) and minimum segment lengths of 10 (top right), 30 (bottom left) and 40 (bottom right). In each plot we show data (grey dots), true mean (blue dashed line), true changepoints (blue vertical dashed lines), estimated mean (black line) and estimated changepoints (red vertical liens)}
%\label{fig:cpop-minseg}
%\end{figure}
%
%
\begin{figure}
\centering
{\includegraphics[scale=.5]{figures/cpop_minseg_ggplot.pdf}}
\caption{Results of analysing data with $t_4$ noise with no minimum segment length (top left) and minimum segment lengths of 10 (top right), 30 (bottom left) and 40 (bottom right). In each plot we show data (grey dots), true mean (blue dashed line), true changepoints (blue vertical dashed lines), estimated mean (black line) and estimated changepoints (red vertical liens)}
\label{fig:cpop-minseg}
\end{figure}
%
%
Results from CPOP with different minimum segment lengths are shown in Figure \ref{fig:cpop-minseg}. If we do not impose a minimum segment length, then we estimate 11 changepoints, including three cluster of changes that overfit to the noise. By imposing a minimum segment length of 10 we avoid the over-fitting. For this example, the computational cost of running \code{cpop} with the minimum segment length is about 15 times larger than when we do not assume a minimum segment length.

@@

The idea of CROPS is that it allows a penalised cost method to be implemented for all penalty values in an interval. This is implemented within the \pkg{cpop} package by the function:
\begin{CodeInput}
cpop.crops(y, x = 1:length(y), grid = x, beta_min = 1.5 * log(length(y)),
  beta_max = 2.5 * log(length(y)), sd =  sqrt(mean(diff(diff(y))^2)/6), 
  minseglen = 0, prune.approx = FALSE)
\end{CodeInput}
The arguments of \code{cpop.crops} are identical to those of \code{cpop} except that, rather than specifying a single penalty value (\code{beta}), the range of penalty values to be used is specified by \code{beta_min} and \code{beta_max}, which fix the smallest and largest penalty value to be used. The output is an instance of an S4 class that contains details of all segmentations found by minimising the penalised cost for some penalty value in the interval between \code{beta_min} and \code{beta_max}.

@@

We simulated data as per Section \ref{sec:cpop} but with \code{sd = 1.5}. To run CPOP for a range of $\beta\in[7.5, 50]$ we use
\begin{CodeInput}
R> res.crops <- cpop.crops(y, x, beta_min = 5, beta_max = 50, sd = 1)
\end{CodeInput}
For our example $2 \log n = 10.6$, so this is equivalent to trying noise standard deviation in the range $[0.8,2.2]$.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> res.crops <- cpop.crops(y, x, beta_min = 8, beta_max = 200, sd = 1)
R> segs <- segmentations(res.crops)
R> p <- ggplot(data = segs, aes(x = m))
R> p <- p + geom_line(aes(y = Qm))
R> p <- p + geom_vline(xintercept = 10, color = "red")
R> p <- p + xlab("No. of changepoints") + ylab("unpenalised cost")
R> plot(p)
\end{CodeInput}
\end{CodeChunk}
This runs \code{cpop.crops} and then uses the fact that the output of \code{segmentations} includes columns that give the number of changepoints and the unpenalised cost of each segmentation. These columns are labelled \code{"m"} and \code{"Qm"} respectively.
The plot gives a clear elbow, see Figure \ref{fig:crops.cpopB}, and this corresponds to a correct estimate of the number of changes. %The fitted changes and estimated mean function are also shown in Figure \ref{fig:crops.cpopB}.
%
%\begin{figure}
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-cropsB1.pdf}}
%\end{minipage}%
%\begin{minipage}{.5\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-cropsB2.pdf}}
%\end{minipage}%
%\newline
%\begin{minipage}{1.0\linewidth}
%\centering
%{\includegraphics[scale=.5]{figures/cpop-cropsB3.pdf}}
%\end{minipage}%
%\caption{Correlated noise example. Output from \code{cpop.crops} (top left), unpenalised cost against number of changepoint (rop right) and estimate from segmentation corresponding to "elbow" (bottom).}
%\label{fig:crops.cpopB}
%\end{figure}
%
\begin{figure}
\centering
{\includegraphics{figures/cpop_crops_ggplot.pdf}}
\caption{Correlated noise example. Output from \code{cpop.crops} (top left), unpenalised cost against number of changepoint (top right) and estimate from segmentation corresponding to "elbow" (bottom).}
\label{fig:crops.cpopB}
\end{figure}

@@

R> res <- cpop(y, x, grid, sd = sqrt(sig2), minseglen = 0.09, beta = 2*log(200))
\end{CodeInput}
\end{CodeChunk}
Here we estimate the noise variance, \code{sig2}, based on the variance of the double difference of the data. For regions where the mean is linear and the data is evenly spaced, taking the double difference will lead to a mean zero process. If the noise is IID with variance $\sigma^2$ then the double-differenced process will have a marginal variance that is $6\sigma^2$. Thus our estimate is the empirical mean of the square of the double-difference data divided by 6. The original data is evenly spaced in in terms of wavenumber, but as we take logs, \code{x} is unevenly spaced: so this estimator will be biased in our setting. However it will give a reasonable ball-park figure for an initial run of \code{cpop} -- the residuals from which can then be used to get a better estimate of the noise variance. We use a evenly spaced grid for possible change-point locations. To avoid the potential for adding multiple changepoints between two observations, we set a minimum segment length of 0.09 (as the largest distance between consecutive \code{x} values is 0.08).

@@

The output is shown in the top-right plot of Figure \ref{fig:crops.data}, and appears to be over-fitting to the early part of the series. This is because the noise variance is heterogeneous, and decreasing with \code{x}. However, given our initial fit we can use the residuals to estimate the noise variance. The noise for the spectra is expected to be approximately inversely proportional to the wavenumber. By using a Taylor-expansion, we have the variance of the noise for the log-spectra should be approximately the variance of the noise of the spectra divided by the square of the mean of the spectra. As the mean of the spectra is roughly a power of the wavenumber, this suggests using a model for the variance, $\sigma^2_x$ say, depending on $x$ as $\log \sigma^2_x = a + bx$: so that the variance is proportional to some power of the wavenumber. We can estimate the parameters of this model by maximising the log-likelihood of Gaussian model for the residuals with this form for the variance.
\begin{CodeChunk}
\begin{CodeInput}
R> r2 <- residuals(res)^2
R> loglik <- function(par)
R> { 
R>  return(length(r2) * par[1] + par[2] * sum(x) 
  + sum(r2 / (exp(par[1]+par[2]*x))))
R> }
R> est.hat <- optim(c(0,0), loglik)
R> sig2 <- exp(est.hat$par[1] + est.hat$par[2]*x)
R> res2 <- cpop(y, x, grid, sd = sqrt(sig2), 
+               minseglen = 0.09, beta =2*log(200))
\end{CodeInput}
\end{CodeChunk}
Here we have calculated the maximum likelihood estimates by using \code{optim} to minimise minus the log-likelihood. The resulting output from \code{cpop} is shown in the bottom left plot of Figure \ref{fig:crops.data}.  The first two changes could represent real regime transitions relating to the inviscid fluid physics that one would see in the real ocean \cite[see Figure 6a of][]{callies2013interpreting}, while the three changes for the largest values of \code{x} may relate to a breakdown in the numerical ocean model near the highest wavenumber of the ocean model grid \cite[]{soufflet2016effective}. The estimates of the spectra for all four series, obtained by repeating this approach, is also shown in Figure \ref{fig:crops.data}. For this application, the residuals from the fitted model appear to be uncorrelated and sub-Gaussian, so using the fit based on the default penalty choice is reasonable. Though one could also explore segmentations for other penalty choices using \code{cpop.crops} as in the previous section.
